---
title: "Mechanistic Model Usage"
author: "Kayla Zhang, Damie Pak, Megan Greischar"
date: "2025-10-20
output:
  bookdown::pdf_document2:
    toc: no
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Our packages for ODE modeling and plotting:
```{r}
require(deSolve)
require(ggplot2)
require(gridExtra)
require(grid)
```

Here, we have our ODE models, the first two of which are based on the 
within-host model from King *et al*., 2009, where immune cells are assumed to
kill pathogens; the first is unmodified, while the second imposes a carrying 
capacity on pathogen load. The third assumes that immune upregulation saturates
as pathogen load increases, and that instead of killing, the immune response 
instead slows pathogen replication. 
```{r}
within.host.model <- function(t, x, params) {
  P <- x[1]
  X <- x[2]

  with(
    as.list(params),
    {
      dP <- r*P - k*X*P
      dX <- a - d*X + y*k*X*P
    
      res <- c(dP, dX)
      list(res)
    }
  ) 
}

within.host.model.cap <- function(t, x, params) {
  P <- x[1]
  X <- x[2]

  with(
    as.list(params),
    {
      dP <- r*P*(1 - P/(5*10^5)) - k*X*P
      dX <- a - d*X + y*k*X*P
    
      res <- c(dP, dX)
      list(res)
    }
  ) 
}

sat.imm.model <- function(t, x, params) {
  P <- x[1]
  X <- x[2]
  
  with(
    as.list(params),
    {
      dP <- r*P/(1+b*X) - dp*P
      dX <- alpha - d*X + gx*P*(Xmax - X)/(P+hp)
      
      res <- c(dP, dX)
      list(res)
    }
  ) 
}
```

Below, we define some useful functions for setting parameters and creating 
dataframes containing the simulated data from our model. We focus on varying 
pathogen replication rate in the present study, so we maintain constant values 
for the remaining parameters based on values used in King *et al*., 2009 (except
for immune cell kill rate of the pathogen, which varies based on the within-host 
model to more easily show both pre- and post-symptomatic transmission later on). 
We truncate the dataframes when pathogen load reaches 0, assuming pathogen clearance. 
```{r}
desolve_function <- function(tstep, r, k, model, id) {
  time <- seq(0,2, by=tstep)

  params <- c(y = 1e-4, k = k, r = r, a = 1, d = 0.5)
  
  df <- data.frame(lsoda(c(P=id, X=0), time, model, 
                       params))
  
  if(any(df$P <= 0)){
    clearance <- min(which(df$P <= 0))
    df[clearance:length(time),] <- NA
  }
  
  df <- data.frame(time=time, 
                  pathogen=df$P, 
                  immune=df$X, 
                  r=r)
  
  return(df)
}

desolve_function.satimm <- function(tstep, r, model, id) {
  time <- seq(0,2, by=tstep)

  params <- c(alpha=1, d=1, b=1e-1, dp=5, gx=1e-12, hp=1e6*10, 
              Xmax=(10)*5e14, r=r)
  
  df <- data.frame(lsoda(c(P=id, X=0), time, model, 
                       params))
  
  if(any(df$P <= 0)){
    clearance <- min(which(df$P <= 0))
    df[clearance:length(time),] <- NA
  }
  
  df <- data.frame(time=time, 
                  pathogen=df$P, 
                  immune=df$X, 
                  r=r)
  
  return(df)
}
```

Here we provide functions for calculating time of peak pathogen load, symptom
onset and end, as well as the delay between symptom 
onset and peak pathogen load. If time of peak pathogen load comes before time of 
symptom onset, meaning the calculated delay is negative, then by our definition, 
there is pre-symptomatic transmission. 
```{r}
peak.pathogenloadtime <- function(df) {
  peak.pl <- max(df$pathogen, na.rm=TRUE)
  return(df[which(df$pathogen==peak.pl),]$time)
}

symptom.onset <- function(df, s.compartment, thresh) {
  symptomatic <- df[which(df[s.compartment] >= thresh),]$time
  if(length(symptomatic) > 0) {
    return(min(symptomatic))
  } else {
    return(NA)
  }
}

symptom.end <- function(df, s.compartment, thresh) {
  symptomatic <- df[which(df[s.compartment] >= thresh),]$time
  if(length(symptomatic) > 0) {
    return(max(symptomatic))
  } else {
    return(NA)
  }
}

delay.calculation <- function(df, symp.thresh) {
  return(peak.pathogenloadtime(df) - symptom.onset(df, 'immune', symp.thresh))
}
```

The first model we examine is one in which symptoms are dependent only on 
pathogen load. This is represented by a dotted horizontal red line depicting the 
threshold for symptoms; if pathogen load falls above that threshold, then symptoms 
are present. We have defined pre-symptomatic transmission as reaching peak pathogen 
load before symptom onset. This is a conservative method of estimating 
pre-symptomatic transmission. 

The second model we examine is one in which symptoms are
dependent only on immune response. When immune response rises above a certain 
threshold, the patient is symptomatic.

Maintaining the assumption that symptoms are dependent only on the immune
response, the third model is one in which immune cells inhibit pathogen replication, 
rather than directly killing them.

Lastly, still maintaining the immune response symptom threshold, the fourth is 
one in which immune cells directly kill pathogens, but there is
a carrying capacity in place for pathogen load.

We produce simulations of these four models to demonstrate their dynamics.
```{r, fig.height = 7.5, fig.width = 6.5}
plot.model.simulations <- function(symp.thresh, data, 
                                    cex, label,
                                    xlimlow, xlimhigh, s.compartment) {
  
  col <- ifelse(s.compartment=='pathogen', 'blue', 'red')
  
  time <- data$time
  path.load <- data$pathogen
  immune <- data$immune
  peak.load.time <- peak.pathogenloadtime(data)
  
  if (s.compartment == 'pathogen') {
    plot(time, path.load, col=col, xlab="", ylab="", xlim=c(xlimlow, xlimhigh), log='y',
         ylim=10^c(-1, 10), cex.lab=cex, cex.axis=cex, mgp=c(0, 1, 0), 
         yaxt='n', tck=-0.04, type="l", lwd=3, bty='l')
    points(peak.load.time, path.load[which(time==peak.load.time,)],
         cex=1.5, pch=16, col=col)
    lines(time, immune, col=gray(0.25), type="l", lty=2)
    axis(2, cex.axis=cex, at=10^(seq(-1, 9, by=2)), 
     labels=c(expression(10^-1), 
              expression(10^1),
              expression(10^3),
              expression(10^5),
              expression(10^7),
              expression(10^9)), las=2, mgp=c(0, 1, 0))
  } else {
    plot(time, immune, col=col, xlab="", ylab="", xlim=c(xlimlow, xlimhigh), log='y',
         ylim=10^c(-1, 10), cex.lab=cex, cex.axis=cex, mgp=c(0, 1, 0), 
         yaxt='n', lty = 2, tck = -0.04, type="l", lwd=3, bty='l')
    points(peak.load.time, path.load[which(time==peak.load.time,)],
         cex=1.5, pch=16)
    lines(time, path.load, col=gray(0.25), type="l")
    axis(2, cex.axis=cex, at=10^(seq(-1, 9, by=2)), 
     labels=c(expression(10^-1), 
              expression(10^1),
              expression(10^3),
              expression(10^5),
              expression(10^7),
              expression(10^9)), las=2, mgp=c(0, 1, 0))
  }

  abline(h=symp.thresh, lty=3, col=col, lwd=2)
  
  polygon(x=c(symptom.onset(data, 
                            s.compartment,
                            symp.thresh),
              symptom.onset(data, 
                            s.compartment,
                            symp.thresh),
              symptom.end(data, 
                            s.compartment,
                            symp.thresh), 
              symptom.end(data, 
                            s.compartment,
                            symp.thresh)), 
          y=10^c(-5, 10, 10, -5), 
          col=rgb(200, 200, 200, alpha=100,
                  maxColorValue=255),
          border = NA)
  
  text(0.01 * (xlimhigh - xlimlow) + xlimlow, 10^(.95 * (9.5)) + 10^-1, 
       labels=label, cex=cex, adj=0)
}

path.symp.thresh <- 10^6
immune.sympthresh <- 15
satimm.sympthresh <- 90
immune.pathcap.sympthresh <- 2

lowrep <- 30
highrep <- 300

data.lowreplication.nocap <- desolve_function(tstep=10^-4, 
                            k=3.5, 
                            r=lowrep,
                            model=within.host.model, 
                            id=10^4)
data.highreplication.nocap <- desolve_function(tstep=10^-4, 
                             k=3.5, 
                             r=highrep,
                             model=within.host.model, id=10^4) 

data.lowreplication.satimm <- desolve_function.satimm(tstep=10^-4,
                                                            r=lowrep,
                                                            model=sat.imm.model,
                                                            id=10^4)

data.highreplication.satimm <- desolve_function.satimm(tstep=10^-4,
                                                            r=highrep,
                                                            model=sat.imm.model,
                                                            id=10^4)

data.lowreplication.cap <- desolve_function(tstep=10^-4, 
                            k=7, r=lowrep,
                            model=within.host.model.cap, id=10^4)
data.highreplication.cap <- desolve_function(tstep=10^-4, 
                             k=7, r=highrep,
                             model=within.host.model.cap, id=10^4) 

# Adjust time display to sensible numbers
data.lowreplication.nocap$time <- data.lowreplication.nocap$time * 7
data.highreplication.nocap$time <- data.highreplication.nocap$time * 7
data.lowreplication.satimm$time <- data.lowreplication.satimm$time * 7
data.highreplication.satimm$time <- data.highreplication.satimm$time * 7
data.lowreplication.cap$time <- data.lowreplication.cap$time * 7
data.highreplication.cap$time <- data.highreplication.cap$time * 7

cex <- 1.25
cex_labels <- 0.9

tiff('Model schematics.tiff', res=500, width=6.5, height=7.5, unit='in')
{
  par(mfrow=c(4, 2), oma=c(0, 2.5, 1.5, 0), mai=c(0.42, 0.32, 0.3, 0.42))
    
  plot.model.simulations(symp.thresh = path.symp.thresh,
                         data = data.lowreplication.nocap, 
                         cex = cex, 
                         label = "Asymptomatic", 
                         xlimlow = 0, xlimhigh = 2.5, 
                         s.compartment = 'pathogen')
  
  mtext(text="SLOW REPLICATION", font=2, side=3, line=2.5, cex=cex_labels)
  
  mtext(text="A. Base model", side=3, line=1, 
        cex=cex_labels, adj=0)
  
  plot.model.simulations(symp.thresh = path.symp.thresh, 
                         data = data.highreplication.nocap, 
                         cex = cex, 
                         label = "Post-symptomatic", 
                         xlimlow = 0, xlimhigh = 0.3, 
                         s.compartment = 'pathogen')
  
  mtext('Pathogen symptom threshold', side=4, line=2, cex=0.9, col='blue')

  # Brackets to indicate that shaded regions = symptomatic

  usr <- par("usr")
  
  par(xpd=NA)
  bracket_y <- grconvertY(1.05, from="npc", to="user")
  bracket_tick_length <- grconvertY(1.05, from="npc", to="user") -
    grconvertY(1.02, from = "npc", to = "user")

  segments(symptom.onset(data.highreplication.nocap, 'pathogen', path.symp.thresh), 
           bracket_y + bracket_tick_length, 
           symptom.end(data.highreplication.nocap, 'pathogen', path.symp.thresh), 
           bracket_y + bracket_tick_length)
  
  segments(symptom.onset(data.highreplication.nocap, 'pathogen', path.symp.thresh), 
           bracket_y + bracket_tick_length, 
           symptom.onset(data.highreplication.nocap, 'pathogen', path.symp.thresh), 
           bracket_y)
  
  segments(symptom.end(data.highreplication.nocap, 'pathogen', path.symp.thresh), 
           bracket_y + bracket_tick_length,
           symptom.end(data.highreplication.nocap, 'pathogen', path.symp.thresh),
          bracket_y)
  
  text((symptom.onset(data.highreplication.nocap, 'pathogen', path.symp.thresh) +
        symptom.end(data.highreplication.nocap, 'pathogen', path.symp.thresh)) / 2, 
        grconvertY(1.15, from="npc", to="user"),
        labels='Symptomatic')
  
  par(xpd=FALSE)
  
  mtext(text="FAST REPLICATION", font=2, side=3, line=2.5, cex=cex_labels)
  
  plot.model.simulations(symp.thresh = immune.sympthresh, 
                         data = data.lowreplication.nocap, 
                         cex = cex, 
                         label = "Pre-symptomatic", 
                         xlimlow = 0, xlimhigh = 2.5, 
                         s.compartment = 'immune')
  
  mtext("Abundance", side=2, line=1, outer=TRUE, adj = 0.5075, cex=0.9)
  
  mtext(text="B. Base model", 
        side=3, line=1, cex=cex_labels, adj=0)
  
  plot.model.simulations(symp.thresh = immune.sympthresh, 
                         data = data.highreplication.nocap, 
                         cex = cex, 
                         label = "Post-symptomatic", 
                         xlimlow = 0, xlimhigh = 0.3, 
                         s.compartment = 'immune')
  
  plot.model.simulations(symp.thresh = satimm.sympthresh,
                         data = data.lowreplication.satimm, 
                         cex = cex, 
                         label = "Pre-symptomatic", 
                         xlimlow = 0, xlimhigh = 8, 
                         s.compartment = 'immune')
  
  mtext(text="C. Immune inhibition of pathogen replication", 
        side=3, line=1, cex=cex_labels, adj=0)
  
  plot.model.simulations(symp.thresh = satimm.sympthresh, 
                         data = data.highreplication.satimm, 
                         cex = cex, 
                         label = "Post-symptomatic", 
                         xlimlow = 0, xlimhigh = 4, 
                         s.compartment = 'immune')
  
  mtext('Immune effector symptom threshold', side=4, line=2, cex=0.9,col='red')
  
  plot.model.simulations(symp.thresh = immune.pathcap.sympthresh, 
                         data = data.lowreplication.cap, 
                         cex = cex, 
                         label = "Post-symptomatic",
                         xlimlow = 0, xlimhigh = 3, 
                         s.compartment = 'immune')
  
  mtext(text="D. Pathogen carrying capacity", 
        side=3, line=1, cex=cex_labels, adj=0)
  
  plot.model.simulations(symp.thresh = immune.pathcap.sympthresh, 
                         data = data.highreplication.cap, 
                         cex = cex, 
                         label = "Pre-symptomatic", 
                         xlimlow = 0, xlimhigh = 1, 
                         s.compartment = 'immune')
  
  legend(0.53, 10^10, legend=c('Peak pathogen load', 
                             'No symp/symp',
                             'Symptom threshold',
                             'Immune effector', 
                             'Pathogen'
                           ), 
         lty=c(NA, NA, 3, 2, 1), pch=c(16, NA, NA, NA, NA), bty="n",
         xpd=TRUE, cex=0.9, col=c('black', NA, 'black', 'blue', 'red'))
  
  mtext("Time (weeks)", side=1, line=-1, outer=TRUE, adj=0.48, cex=0.9)
}

dev.off()
```

Now, we vary pathogen replication rate (r) and plot the delay between
symptom onset and peak pathogen load against various r values for all models to
investigate the correlation between replication rate and
pre-symptomatic transmission. 
```{r, fig.height = 6.5, fig.width = 3.5}
get.delays.df <- function(data, param.vals, symp.thresh) {
  
  delays.df <- do.call(rbind, lapply(param.vals, function(x) 
    data.frame(r=x,
               delay.symponset.peakload=
                 delay.calculation(data[data$r==x,], symp.thresh),
               symp.thresh=symp.thresh))) 
  
  return(delays.df)
}

get.peak.imm <- function(data) {
  
  peak <- peak.pathogenloadtime(data)
  
  imm.at.peak <- data[data$time==peak,]$immune
  
  return(imm.at.peak)
  
}

get.peak.imm.df <- function(data, param.vals) {
  
  peaks <- do.call(rbind, lapply(param.vals, function(x) 
    data.frame(r=x,
               imm.at.peak=get.peak.imm(data[data$r==x,])))) 
  
  return(peaks)
  
}

r_vals <- seq(10, 300, by=10)

varyingr.nocap <- data.frame(k=3.5, r=r_vals, dose=10^4)

varyingr.nocap <- do.call(rbind, mapply(desolve_function, 
                                  tstep=10^-4, 
                                  r=varyingr.nocap$r,
                                  k=varyingr.nocap$k,
                                  id=varyingr.nocap$dose,                                  
                                  MoreArgs=list(model=within.host.model),
                                  SIMPLIFY = FALSE))

varyingr.satimm <- data.frame(r=r_vals, dose=10^4)
varyingr.satimm <- do.call(rbind, mapply(desolve_function.satimm, 
                                  tstep=10^-4, 
                                  r=varyingr.satimm$r, 
                                  id=varyingr.satimm$dose,                                  
                                  MoreArgs=list(model=sat.imm.model),
                                  SIMPLIFY = FALSE))

varyingr.cap <- data.frame(k= 7, r=r_vals, dose=10^4)

varyingr.cap <- do.call(rbind, mapply(desolve_function, 
                                  tstep=10^-4, 
                                  r=varyingr.cap$r,
                                  k=varyingr.cap$k,
                                  id=varyingr.cap$dose,                                  
                                  MoreArgs=list(model=within.host.model.cap),
                                  SIMPLIFY = FALSE))

delays.byr.lowsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=10)

delays.byr.medsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=15)

delays.byr.highsympthresh.nocap <- get.delays.df(data=varyingr.nocap, 
                                          param.vals=r_vals,
                                          symp.thresh=20)

delays.byr.lowsympthresh.satimm <- get.delays.df(data=varyingr.satimm, 
                                          param.vals=r_vals,
                                          symp.thresh=70)

delays.byr.medsympthresh.satimm <- get.delays.df(data=varyingr.satimm, 
                                          param.vals=r_vals,
                                          symp.thresh=90)

delays.byr.highsympthresh.satimm <- get.delays.df(data=varyingr.satimm, 
                                          param.vals=r_vals,
                                          symp.thresh=110)

delays.byr.lowsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=1)

delays.byr.medsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=2)

delays.byr.highsympthresh.cap <- get.delays.df(data=varyingr.cap, 
                                          param.vals=r_vals,
                                          symp.thresh=3)

immpeak.byr.base <- get.peak.imm.df(data=varyingr.nocap, 
                                                  r_vals)

plot.varyingr <- function(delaysbyr1, delaysbyr2, delaysbyr3, 
                          highlighted_points, size, ylow, yhigh,
                          xlow, xhigh, legendposx, legendposy) {
  
  return(ggplot(delaysbyr1, aes(x=r, y=delay.symponset.peakload, 
                         color = as.character(symp.thresh))) + theme_classic() +
    theme(
          plot.title = element_text(hjust = 0.5, vjust = 3, size=size,
                                    face="bold"), 
          plot.margin = unit(c(0.05, 0, 0, -0.05),"in"),
          axis.title.y = element_text(margin = margin(r=4), size=size),
          axis.title.x = element_text(margin = margin(t=8), size=size),
          axis.text.x = element_text(size=size, margin = margin(t=2)),
          axis.text.y = element_text(size=size, margin = margin(r=2)),
          axis.ticks.length=unit(0.05, "in"),
          legend.key.size = unit(0.001, "in"),
          legend.text = element_text(size=size * .9),
          legend.title = element_text(size=size * .9),
          legend.position = c(legendposx, legendposy)) +
      ylim(ylow, yhigh) + 
    xlim(xlow, xhigh) +
  geom_point(size=1) + 
    geom_line() + geom_point(data=delaysbyr2, size=1) + 
    geom_line(data=delaysbyr2) + geom_point(data=delaysbyr3, size=1) +
    geom_line(data=delaysbyr3) +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
    geom_point(data=highlighted_points, aes(x=r, y=delay.symponset.peakload), color="red", size=1)) 
}

cols <- colorRampPalette(c("deepskyblue", "deepskyblue4"))(3)

tiff('Delays varying r.tiff', res=500, width=3.5, height=5.5, unit='in')
{plot.varyingr.nocap <- plot.varyingr(delays.byr.lowsympthresh.nocap, 
              delays.byr.medsympthresh.nocap, 
              delays.byr.highsympthresh.nocap, 
              highlighted_points=delays.byr.medsympthresh.nocap
              [delays.byr.medsympthresh.nocap$r==lowrep | 
                  delays.byr.medsympthresh.nocap$r==highrep,],
              size=8.5, 
              ylow=-.06, 
              yhigh=.005, 
              xlow=0, 
              xhigh=400,
              legendposx=0.8,
              legendposy=0.3) + 
  ggtitle("A) Base model") + 
  xlab("") +
  ylab("") +
  geom_text(aes(x=360, y=-0.003, label = "Pre-symptomatic"),
                color="black", size=2.25) +
  geom_text(aes(x=360, y=0.0035, label = "Post-symptomatic"),
                color="black", size=2.25) +
  scale_color_manual(name="Symptom Threshold", values = c("10"= cols[1], 
                                                          "15"= cols[2], 
                                                          "20"= cols[3]),
                     limits=c("10", "15", "20"))
  
plot.varyingr.satimm <- plot.varyingr(delays.byr.lowsympthresh.satimm, 
              delays.byr.medsympthresh.satimm, 
              delays.byr.highsympthresh.satimm, 
              highlighted_points=delays.byr.medsympthresh.satimm
              [delays.byr.medsympthresh.satimm$r==lowrep | 
                  delays.byr.medsympthresh.satimm$r==highrep,],
              size=8.5,
              ylow=-.45, 
              yhigh=0.15, 
              xlow=0, 
              xhigh=400,
              legendposx=0.8,
              legendposy=0.3) + 
  ggtitle("B) Immune inhibition of pathogen replication") + 
  xlab("") +
  ylab("") +
  geom_text(aes(x=360, y=-0.03, label = "Pre-symptomatic"),
                color="black", size=2.25) +
  geom_text(aes(x=360, y=0.03, label = "Post-symptomatic"),
                color="black", size=2.25) +
  scale_color_manual(name="Symptom Threshold", values = c("70"= cols[1], 
                                                          "90"= cols[2], 
                                                          "110"= cols[3]),
                     limits=c("70", "90", "110"))  

plot.varyingr.cap <- plot.varyingr(delays.byr.lowsympthresh.cap,
              delays.byr.medsympthresh.cap, 
              delays.byr.highsympthresh.cap, 
              highlighted_points=delays.byr.medsympthresh.cap
              [delays.byr.medsympthresh.cap$r==lowrep | 
                  delays.byr.medsympthresh.cap$r==highrep,],
              size=8.5, 
              ylow=-.04, 
              yhigh=.02, 
              xlow=0, 
              xhigh=400,
              legendposx=0.8,
              legendposy=0.3) + 
  ggtitle("C) Pathogen carrying capacity") +
  xlab("Replication rate (r)") +
  ylab("") +
  geom_text(aes(x=360, y=-0.003, label = "Pre-symptomatic"),
                color="black", size=2.25) +
  geom_text(aes(x=360, y=0.00325, label = "Post-symptomatic"),
                color="black", size=2.25) + 
  scale_color_manual(name="Symptom Threshold", values = c("1"= cols[1], 
                                                          "2"= cols[2], 
                                                          "3"= cols[3]))

  grid.arrange(plot.varyingr.nocap, plot.varyingr.satimm, plot.varyingr.cap,
               nrow=3, 
               left=textGrob("Day of peak viral load \n minus day of symptom onset",
               gp=gpar(fontsize = 8.5),
               rot=90))
}
dev.off()
```

We show that for the range of replication rates used, immune response peaks after 
pathogen load.
```{r, fig.height = 6.5, fig.width = 3.5}
peak.immunetime <- function(df) {
  peak.immune <- max(df$immune, na.rm=TRUE)
  return(df[which(df$immune==peak.immune),]$time)
}

get.peaktimes.df <- function(data, symp.thresh) {
  
  peaktimes.df <- do.call(rbind, lapply(r_vals, function(x) 
    data.frame(r=x,
               pathogen.peaktime=peak.pathogenloadtime(data[data$r==x,]),
               immune.peaktime=peak.immunetime(data[data$r==x,]),
               symp.onset=symptom.onset(data[data$r==x,], 'immune', symp.thresh)))) 
  
  return(peaktimes.df)
}

peaktimes.cap <- get.peaktimes.df(varyingr.cap, 2)
peaktimes.nocap <- get.peaktimes.df(varyingr.nocap, 15)
peaktimes.satimm <- get.peaktimes.df(varyingr.satimm, 90)
cex <- 1.25
cex.mtext <- 0.9

{
  par(mfrow=c(3, 1), mai=c(0.52, 0.62, 0.42, 0.1))
  
  plot(pathogen.peaktime ~ r, data=peaktimes.nocap, ylim=c(0, 0.15), type="l",
       xlab="", ylab="", yaxt="n", bty="l", 
       mgp=c(2.25, 0.75, 0), cex.axis=cex)
  axis(2, las=2, cex.axis=cex)
  lines(immune.peaktime ~ r, data=peaktimes.nocap, col="red", lty=2)
  
  mtext("A) Base model", side=3, line=1.5, cex=cex.mtext)
  
  plot(pathogen.peaktime ~ r, data=peaktimes.satimm, ylim=c(0, 0.9), type="l",
       xlab="", ylab="", yaxt="n", bty="l", cex.axis=cex)
  axis(2, las=2, cex.axis=cex)
  lines(immune.peaktime ~ r, data=peaktimes.satimm, col="red", lty=2)
  mtext("Time of max abundance", 2, line=3.5, cex=cex.mtext)
  
  mtext("B) Immune inhibition of pathogen replication", side=3, line=1.5, cex=cex.mtext)
  
  plot(pathogen.peaktime ~ r, data=peaktimes.cap, ylim=c(0, 0.15), type="l",
       xlab="", ylab="", yaxt="n", bty="l", cex.axis=cex)
  axis(2, las=2, cex.axis=cex)
  lines(immune.peaktime ~ r, data=peaktimes.cap, col="red", lty=2)
  
  mtext("C) Pathogen carrying capacity", side=3, line=1.5, cex=cex.mtext)
  
  legend(125, 0.15, legend=c('Immune effector',
                             'Pathogen'), 
         lty=c(2, 1), col=c("red", "black"), bty="n")
  
  mtext("Replication rate (r)", side=1, line=2.5, cex=cex.mtext)
}
```

We also compare the distance between symptom onset and peak pathogen load 
at different replication rates.
```{r, fig.height = 6.5, fig.width = 3.5}
{
  par(mfrow=c(3, 1), mai=c(0.52, 0.62, 0.42, 0.1))
  
  plot(pathogen.peaktime ~ r, data=peaktimes.nocap, ylim=c(0.02, 0.2), type="l",
       xlab="", ylab="", yaxt="n", bty="l", 
       mgp=c(2.25, 0.75, 0), cex.axis=cex)
  axis(2, las=2, cex.axis=cex)
  lines(symp.onset ~ r, data=peaktimes.nocap, col="red", lty=2)
  
  mtext("A) Base model", side=3, line=1.5, cex=cex.mtext)
  
  plot(pathogen.peaktime ~ r, data=peaktimes.satimm, ylim=c(0, 0.6), type="l",
       xlab="", ylab="", yaxt="n", bty="l", cex.axis=cex)
  axis(2, las=2, cex.axis=cex, at=seq(0, 0.6, length=4))
  lines(symp.onset ~ r, data=peaktimes.satimm, col="red", lty=2)
  mtext("Time", 2, line=3.5, cex=cex.mtext)
  
  mtext("B) Immune inhibition of pathogen replication", side=3, line=1.5, cex=cex.mtext)
  
  plot(pathogen.peaktime ~ r, data=peaktimes.cap, ylim=c(0.02, 0.2), type="l",
       xlab="", ylab="", yaxt="n", bty="l", cex.axis=cex)
  axis(2, las=2, cex.axis=cex)
  lines(symp.onset ~ r, data=peaktimes.cap, col="red", lty=2)
  
  mtext("C) Pathogen carrying capacity", side=3, line=1.5, cex=cex.mtext)
  
  legend(90, 0.15, legend=c('Symptom onset',
                             'Pathogen load'), 
         lty=c(2, 1), col=c("red", "black"), bty="n")
  
  mtext("Replication rate (r)", side=1, line=2.5, cex=cex.mtext)
}

```